Text embedding models yield high-resolution insights into conceptual knowledge from short multiple-choice quizzes

### 2024-12-23 paper/main.tex
rewrote fig 6 results section

If we are able to accurately estimate a participant's knowledge about the
content tested by a given question, our estimates of their knowledge should
carry some predictive information about whether they are likely to answer that
question correctly or incorrectly. We developed a statistical approach to test
this claim. For each quiz question a participant answered, in turn, we used
Equation~\ref{eqn:prop} to estimate their knowledge at the given question's
embedding space coordinate based on other questions that participant answered
on the same quiz. We repeated this for all participants, and for each of the
three quizzes. Then, separately for each quiz, we fit a generalized linear
mixed model (GLMM) with a logistic link function to explain the likelihood of
correctly answering a question as a function of estimated knowledge for its
embedding coordinate, while accounting for random variation among participants
and questions (see \nameref{subsec:glmm}). To assess the predictive value of
the knowledge estimates, we compared each GLMM to an analogous (i.e., nested)
``null'' model that did not consider estimated knowledge using parametric
bootstrap likelihood-ratio tests.

@@

We carried out three different versions of the analyses described above,
wherein we considered different sources of information in our estimates of
participants' knowledge for each quiz question. First, we estimated knowledge
at each question's embedding coordinate using \textit{all} other questions
answered by the same participant on the same quiz (``All questions'';
Fig.~\ref{fig:predictions}, top row). This test was intended to assess the
overall predictive power of our approach. Second, we estimated knowledge for
each question about a given lecture using only the other questions (from the
same participant and quiz) about that \textit{same} lecture
(``Within-lecture''; Fig.~\ref{fig:predictions}, middle rows). This test was
intended to assess the \textit{specificity} of our approach by asking whether
our predictions could distinguish between questions about different content
covered by the same lecture. Third, we estimated knowledge for each question
about one lecture using only questions (from the same participant and quiz)
about the \textit{other} lecture (``Across-lecture'';
Fig.~\ref{fig:predictions}, bottom rows). This test was intended to assess the
\textit{generalizability} of our approach by asking whether our predictions
held across the content areas of the two lectures.

In performing these analyses, our null hypothesis is that the knowledge
estimates we compute based on the quiz questions' embedding coordinates do
\textit{not} provide useful information about participants' abilities to answer
those questions. What result might we expect to see if this is the case? To
gain an intuition for this scenario, consider the expected outcome if we
carried out these same analyses using a simple proportion-correct measure in
lieu of our knowledge estimates. Suppose a participant correctly answered $n$
out of $q$ questions on a given quiz. If we hold out a single
\textit{correctly} answered question, the proportion of remaining questions
answered correctly would be $\frac{n - 1}{q - 1}$. Whereas if we hold out a
single \textit{incorrectly} answered question, the proportion of remaining
questions answered correctly would be $\frac{n}{q - 1}$. In this way, the
proportion of correctly answered remaining questions is always \textit{lower}
when the held-out question was answered correctly than when it was answered
incorrectly. Because our knowledge estimates are computed as a weighted version
of this same proportion-correct score (where each held-in question's weight
reflects its embedding-space distance from the held-out question; see
Eqn.~\ref{eqn:prop}), if these weights are uninformative (e.g., randomly
distributed), then we should expect to see this same inverse relationship
between estimated knowledge and performance, on average. On the other hand, if
the spatial relationships among the quiz questions' embeddings \textit{are}
predictive of participants' knowledge about the questions' content, then we
would expect \textit{higher} estimated knowledge for held-out correctly (versus
incorrectly) answered questions.

Before presenting our results, it is worth considering three possible
explanations of why a participant might answer a given question correctly or
incorrectly. One possibility is that the participant simply \textit{guessed}
the answer. A second is that they selected the incorrect answer by mistake,
despite ``knowing'' the correct answer (or vice versa). In both of these
scenarios, the participant's knowledge about the question's content should be
uninformative about their observed response. A third possibility is that the
participant's response reflects their \textit{actual} knowledge about the
question's content. In this case, we \textit{might} expect to see a positive
relationship between the participant's knowledge and their likelihood of
answering the question correctly. However, in order to see this positive
relationship, the participant's knowledge must be structured in a way that is
reflected (at least partially) by the embedding space. In other words, if the
participant's performance reflects their true knowledge, but our text embedding
space does not sufficiently capture the structure of that knowledge, then the
knowledge estimates we generate will not be predictive of the participant's
performance. In the extreme, if the embedding space is completely unstructured
with respect to the content of the quiz questions, then we would expect to see
the negative relationship between estimated knowledge and performance that we
described above.

When we fit a GLMM to estimates of participants' knowledge for each Quiz~1
question based on all other Quiz~1 questions, we observed an outcome consistent
with our null hypothesis: higher estimated knowledge at the embedding
coordinate of a held-out question was associated with a lower likelihood of
answering the question correctly (odds ratio $(OR) = 0.136$, likelihood-ratio
test statistic $(\lambda_{LR}) = 19.749$, 95\%\ $\textnormal{CI} = [14.352,\
26.545],\ p < 0.001$). This outcome suggests that our knowledge estimates do
\textit{not} provide useful information about participants' Quiz~1 performance
when we aggregated across all question content areas. We speculated that this
might either indicate that the knowledge estimates are uninformative in
general, or about Quiz~1 performance in particular. This would be expected, for
example, if participants were guessing about the answers to the Quiz~1
questions (prior to having watched either lecture). When we repeated this
analysis for Quizzes~2 and~3, we found that higher estimated knowledge
for a given question predicted a \textit{greater} likelihood of answering it correctly
(Quiz~2: $OR = 2.905,\ \lambda_{LR} = 17.333,\ 95\%\ \textnormal{CI} =
[14.966,\ 29.309],\ p = 0.001$; Quiz~3: $OR = 3.238,\ \lambda_{LR} = 6.882,\
95\%\ \textnormal{CI} = [6.228,\ 8.184],\ p = 0.016$). Taken together, these
results suggest that our knowledge estimates reliably predict participants'
performance on individual held-out quiz questions, but only after participants
have received at least some training.

We observed a similar pattern of results when used this approach to estimate
participants' knowledge about held-out questions from one lecture using their
performance on other questions from the \textit{same} lecture. Specifically,
for Quiz~1 questions (i.e., prior to watching either), participants' estimated
knowledge for the embedding coordinates of held-out \textit{Four Fundamental
Forces}-related questions estimated using other \textit{Four Fundamental
Forces}-related questions did not reliably predict whether those questions were
answered correctly ($OR = 1.891,\ \lambda_{LR} = 2.293,\ 95\%\ \textnormal{CI}
= [2.091,\ 2.622],\ p = 0.138$). The same was true of knowledge estimates for
held-out \textit{Birth of Stars}-related questions based on other \textit{Birth
of Stars}-related questions ($OR = 0.722,\ \lambda_{LR} = 5.115,\ 95\%\
\textnormal{CI} = [0.094,\ 0.146],\ p = 0.738$). As in our analysis that
included all questions, we speculate that these ``null'' results might reflect
some degree of random guessing on Quiz~1. When we repeated these within-lecture
analyses using questions from Quiz~2 (which participants took immediately after
viewing \textit{Four Fundamental Forces} but prior to viewing \textit{Birth of
Stars}), we found that they now reliably predicted success on \textit{Four
Fundamental Forces}-related questions ($OR = 9.023,\ \lambda_{LR} = 18.707,\
95\%\ \textnormal{CI} = [10.877,\ 22.222],\ p < 0.001$) but not on
\textit{Birth of Stars}-related questions ($OR = 0.306,\ \lambda_{LR} = 5.115,\
95\%\ \textnormal{CI} = [4.624,\ 5.655],\ p = 0.054$). Here, we speculate that
participants might have been guessing about the \textit{Birth of Stars} content
(e.g., prior to having watched it), whereas they might have been drawing on
some structured knowledge about the \textit{Four Fundamental Forces} content
(e.g., from having just watched it). When we applied this approach to Quiz~3
responses (given immediately after viewing \textit{Birth of Stars}), we found
that within-lecture knowledge estimates for \textit{Birth of Stars}-related
questions could now reliably predict success on those questions ($OR = 5.467,\
\lambda_{LR} = 10.670,\ 95\%\ \textnormal{CI} = [7.998, 12.532],\ p = 0.005$).
However, within-lecture knowledge estimates for \textit{Four Fundamental
Forces} questions answered on Quiz~3 were no longer directly related to the
likelihood of successfully answering them and instead exhibited the inverse
relationship we would expect to arise from unstructured knowledge (with respect
to the embedding space; $OR = 0.013,\ \lambda_{LR} = 14.648,\ 95\%\
\textnormal{CI} = [10.695, 23.096],\ p < 0.001$). Speculatively, we suggest
that this may reflect participants forgetting some of the \textit{Four
Fundamental Forces} content (e.g., perhaps in favor of prioritizing encoding
the just-watched \textit{Birth of Stars} content in preparation for the third
quiz). If this forgetting happens in a relatively ``random'' way (with respect
to spatial distance within the embedding space), then it could explain why some
held-out questions about \textit{Four Fundamental Forces} were answered
incorrectly, even if questions at nearby coordinates (i.e., about similar
content) were answered correctly. This might lead our approach to over-estimate
knowledge for held-out questions about ``forgotten'' knowledge that
participants answered incorrectly. Taken together, these within-lecture results
suggest that our approach can distinguish between questions about different
content covered by a single lecture when participants have sufficiently
structured knowledge about its contents, though this specificity may decrease
with time since the relevant material was learned.

Finally, we used this approach to estimate participants' knowledge about
held-out questions from one lecture using their performance on questions from
the \textit{other} lecture. Here we again observed a similar pattern of
results, though with some notable differences. On Quiz~1, we found that
participants' abilities to correctly answer questions about \textit{Four
Fundamental Forces} could be predicted from their responses to questions about
\textit{Birth of Stars} ($OR = 1.896,\ \lambda_{LR} = 7.205,\ 95\%\
\textnormal{CI} = [6.224, 7.524],\ p = 0.038$) and similarly, that their
ability to correctly answer \textit{Birth of Stars}-related questions could be
predicted from their responses to \textit{Four Fundamental Forces}-related
questions ($OR = 1.522,\ \lambda_{LR} = 6.448,\ 95\%\ \textnormal{CI} = [5.656,
6.843],\ p = 0.042$). Given the results from our analyses that included all
questions and within-lecture predictions, we were surprised to find that the
knowledge estimates could reliably (if weakly) predict participants'
performance across content from different lectures. It is possible that this
result reflects a combination of random guessing prior to training (leading to
a weak effect size), alongside some coarse-scale structured knowledge that
participants had about the content prior to watching either lecture. When we
repeated this analysis using questions from Quiz~2, we found participants'
responses to \textit{Four Fundamental Forces}-related questions did
\textit{not} reliably predict their success on \textit{Birth of Stars}-related
questions ($OR = 1.865,\ \lambda_{LR} = 3.205,\ 95\%\ \textnormal{CI} = [3.027,
3.600],\ p = 0.124$), nor did their responses to \textit{Birth of
Stars}-related questions reliably predict their success on \textit{Four
Fundamental Forces}-related questions ($OR = 3.490,\ \lambda_{LR} = 3.266,\
95\%\ \textnormal{CI} = [3.033, 3.866],\ p = 0.093$). These ``prediction
failures'' appear to come from the fact that any signal derived from
participants' knowledge about the content of the \textit{Birth of Stars}
lecture (prior to watching it) is overwhelmed by the much more dramatic increase in
their knowledge about the content of the \textit{Four Fundamental Forces}
(which they watched just prior to taking Quiz~2). This is reflected in their
Quiz~2 performance for questions about each lecture (mean proportion correct
for \textit{Four Fundamental Forces}-related questions on Quiz~2: $0.77$; mean
proportion correct for \textit{Birth of Stars}-related questions on Quiz~2:
$0.36$). When we carried out these across-lecture knowledge predictions using
questions from Quiz~3 (when participants had now viewed \textit{both}
lectures), we could again reliably predict success on questions about both
\textit{Four Fundamental Forces} ($OR = 11.294),\ \lambda_{LR} = 11.055,\ 95\%\
\textnormal{CI} = [9.126, 18.476],\ p = 0.003$) and \textit{Birth of Stars}
($OR = 7.302,\ \lambda_{LR} = 7.068,\ 95\%\ \textnormal{CI} = [6.490, 8.584],\
p = 0.016$) using responses to questions about the other lecture's content.
Across all three versions of these analyses, our results suggest that (by and
large) our knowledge estimates can reliably predict participants' abilities to
answer individual quiz questions, distinguish between questions about similar
content, and generalize across content areas, provided that participants' quiz
responses reflect a minimum level of ``real'' knowledge about both content on
which these predictions are based and that for which they are made.

@@

We projected the resulting 100-dimensional topic vectors (for each second of
video and each quiz question) onto a shared 2-dimensional plane (see
\nameref{subsec:knowledge-maps}). Next, we sampled points from a $100 \times
100$ grid of coordinates that evenly tiled a rectangle enclosing the 2D
projections of the videos and questions. We used
Equation~\ref{eqn:rbf-knowledge} to estimate participants' knowledge at each of
these 10,000 sampled locations, and averaged these estimates across
participants to obtain an estimated average \textit{knowledge map}
(Fig.~\ref{fig:knowledge-maps}A). Intuitively, the knowledge map constructed
from a given quiz's responses provides a visualization of how ``much''
participants knew about any content expressible by the fitted text embedding
model at the point in time when they completed that quiz.

>>>>

If we are able to accurately estimate a participant's knowledge about the
content tested by a given question, our estimates of their knowledge should
carry some predictive information about whether they are likely to answer that
question correctly or incorrectly. We developed a statistical approach to test
this claim. For each quiz question a participant answered, in turn, we used
Equation~\ref{eqn:prop} to estimate their knowledge at the given question's
embedding space coordinate based on other questions that participant answered
on the same quiz. We repeated this for all participants, and for each of the
three quizzes. Then, separately for each quiz, we fit a generalized linear
mixed model (GLMM) with a logistic link function to explain the probability of
correctly answering a question as a function of estimated knowledge for its
embedding coordinate, while accounting for random variation among participants
and questions (see \nameref{subsec:glmm}). To assess the predictive value of
the knowledge estimates, we compared each GLMM to an analogous (i.e., nested)
``null'' model that did not consider estimated knowledge using parametric
bootstrap likelihood-ratio tests.

@@

We carried out three different versions of the analyses described above,
wherein we considered different sources of information in our estimates of
participants' knowledge for each quiz question. First, we estimated knowledge
at each question's embedding coordinate using \textit{all} other questions
answered by the same participant on the same quiz (``All questions'';
Fig.~\ref{fig:predictions}, top row). This test was intended to assess the
overall predictive power of our approach. Second, we estimated knowledge for
each question about a given lecture using only the other questions (from the
same participant and quiz) about that \textit{same} lecture
(``Within-lecture''; Fig.~\ref{fig:predictions}, middle rows). This test was
intended to assess the \textit{specificity} of our approach by asking whether
our predictions could distinguish between questions about different content
covered by the same lecture. Third, we estimated knowledge for each question
about one lecture using only the questions (from the same participant and quiz)
about the \textit{other} lecture (``Across-lecture'';
Fig.~\ref{fig:predictions}, bottom rows). This test was intended to assess the
\textit{generalizability} of our approach by asking whether our predictions 
could extend across the content areas of the two lectures. When computing these 
knowledge estimates, we used a rebalancing procedure to ensure that (for a given 
participant and quiz) the knowledge estimates for correctly and incorrectly 
answered questions were based on the same proportion of correctly answered 
questions (see~\nameref{subsec:glmm}).

%In performing these analyses, our null hypothesis is that the knowledge
%estimates we compute based on the quiz questions' embedding coordinates do
%\textit{not} provide useful information about participants' abilities to answer
%those questions. In other words, our estimates of participants' knowledge
%for questions they answered correctly should (on average) be 
%
%In other words, since our knowledge estimates are computed 
%as a weighted proportion-correct score (where each held-in question's weight reflects
%its embedding-space distance from the held-out question; see
%Eqn.~\ref{eqn:prop}), if these weights are uninformative (e.g., randomly
%distributed), then we should expect that, on average, 
%
%% note that with correction, pcorrect of held in questions is now $\frac{n - 1}{q - 1}$ regardless of held-out question's label
%
%==================================
%
%In performing these analyses, our null hypothesis is that the knowledge
%estimates we compute based on the quiz questions' embedding coordinates do
%\textit{not} provide useful information about participants' abilities to answer
%those questions. What result might we expect to see if this is the case? To
%gain an intuition for this scenario, consider the expected outcome if we
%carried out these same analyses using a simple proportion-correct measure in
%lieu of our knowledge estimates. Suppose a participant correctly answered $n$
%out of $q$ questions on a given quiz. If we hold out a single
%\textit{correctly} answered question, the proportion of remaining questions
%answered correctly would be $\frac{n - 1}{q - 1}$. Whereas if we hold out a
%single \textit{incorrectly} answered question, the proportion of remaining
%questions answered correctly would be $\frac{n}{q - 1}$. In this way, the
%proportion of correctly answered remaining questions is always \textit{lower}
%when the held-out question was answered correctly than when it was answered
%incorrectly. Because our knowledge estimates are computed as a weighted version
%of this same proportion-correct score (where each held-in question's weight
%reflects its embedding-space distance from the held-out question; see
%Eqn.~\ref{eqn:prop}), if these weights are uninformative (e.g., randomly
%distributed), then we should expect to see this same inverse relationship
%between estimated knowledge and performance, on average. On the other hand, if
%the spatial relationships among the quiz questions' embeddings \textit{are}
%predictive of participants' knowledge about the questions' content, then we
%would expect \textit{higher} estimated knowledge for held-out correctly (versus
%incorrectly) answered questions.
%
%Before presenting our results, it is worth considering three possible
%explanations of why a participant might answer a given question correctly or
%incorrectly. One possibility is that the participant simply \textit{guessed}
%the answer. A second is that they selected the incorrect answer by mistake,
%despite ``knowing'' the correct answer (or vice versa). In both of these
%scenarios, the participant's knowledge about the question's content should be
%uninformative about their observed response. A third possibility is that the
%participant's response reflects their \textit{actual} knowledge about the
%question's content. In this case, we \textit{might} expect to see a positive
%relationship between the participant's knowledge and their likelihood of
%answering the question correctly. However, in order to see this positive
%relationship, the participant's knowledge must be structured in a way that is
%reflected (at least partially) by the embedding space. In other words, if the
%participant's performance reflects their true knowledge, but our text embedding
%space does not sufficiently capture the structure of that knowledge, then the
%knowledge estimates we generate will not be predictive of the participant's
%performance. In the extreme, if the embedding space is completely unstructured
%with respect to the content of the quiz questions, then we would expect to see
%the negative relationship between estimated knowledge and performance that we
%described above.

%%\textbf{NOTES FOR REWRITING RESULTS}
%%\begin{itemize}
%%    \item all questions
%%    \begin{itemize}
%%        \item quiz 1: odds ratio $(OR) = 8.126$, 95\%\ $\textnormal{CI} = [3.116,\ 20.123]$, likelihood-ratio test statistic $(\lambda_{LR}) = 17.002,\ p < 0.001$
%%        \item quiz 2: $OR = 14.902,\ 95\%\ \textnormal{CI} = [4.976,\ 39.807],\ \lambda_{LR} = 25.408,\ p < 0.001$
%%        \item quiz 3: $OR = 37.409,\ 95\%\ \textnormal{CI} = [10.425,\ 107.145],\ \lambda_{LR} = 40.948,\ p < 0.001$
%%        \item we can reliably predict success on individual held-out questions regardless of level of knowledge; predictive power grows as participants' knowledge space becomes better defined/structured 
%%    \end{itemize}
%%    \item within-lecture
%%    \begin{itemize}
%%        \item quiz 1, lecture 1: $OR = 15.934,\ 95\%\ \textnormal{CI} = [5.173,\ 38.005],\ \lambda_{LR} = 40.971,\ p < 0.001$
%%        \item quiz 1, lecture 2: $OR = 9.775,\ 95\%\ \textnormal{CI} = [2.93,\ 25.08],\ \lambda_{LR} = 13.924,\ p < 0.001$
%%        \item quiz 2, lecture 1: $OR = 35.126,\ 95\%\ \textnormal{CI} = [5.113,\ 123.868],\ \lambda_{LR} = 32.251,\ p < 0.001$
%%        \item quiz 2, lecture 2: $OR = 4.717,\ 95\%\ \textnormal{CI} = [2.021,\ 9.844],\ \lambda_{LR} = 16.788,\ p < 0.001$
%%        \item quiz 3, lecture 1: $OR = 2.485,\ 95\%\ \textnormal{CI} = [0.724,\ 8.366],\ \lambda_{LR} = 1.984,\ p = 0.170$
%%        \item quiz 3, lecture 2: $OR = 16.902,\ 95\%\ \textnormal{CI} = [3.353,\ 53.265],\ \lambda_{LR} = 23.233,\ p < 0.001$
%%        \item can successfully distinguish between knowledge for questions about different content within a lecture, but may be susceptible to ceiling effects (considering questions included in model, across-participants avg prop correct on quiz3, lecture-1 questions was highest of any set of questions we analyzed (mean: 0.795, std: 0.236)
%%    \end{itemize}
%%    \item across-lecture
%%    \begin{itemize}
%%        \item quiz 1, lecture 1: $OR = 1.896,\ 95\%\ \textnormal{CI} = [0.419,\ 9.088],\ \lambda_{LR} = 0.712,\ p = 0.404$
%%        \item quiz 1, lecture 2: $OR = 1.522,\ 95\%\ \textnormal{CI} = [0.332,\ 6.835],\ \lambda_{LR} = 0.286,\ p = 0.611$
%%        \item quiz 2, lecture 1: $OR = 3.49,\ 95\%\ \textnormal{CI} = [0.739,\ 12.849],\ \lambda_{LR} = 3.266,\ p = 0.083$
%%        \item quiz 2, lecture 2: $OR = 2.199,\ 95\%\ \textnormal{CI} = [0.711,\ 5.623],\ \lambda_{LR} = 2.304,\ p = 0.141$
%%        \item quiz 3, lecture 1: $OR = 11.294,\ 95\%\ \textnormal{CI} = [1.375,\ 47.744],\ \lambda_{LR} = 10.396,\ p < 0.001$
%%        \item quiz 3, lecture 2: $OR = 7.302,\ 95\%\ \textnormal{CI} = [1.077,\ 44.879],\ \lambda_{LR} = 4.708,\ p = 0.038$
%%        \item ability to generalize across content areas depends on participants having structured knowledge about both (A) the content used to estimate knowledge, and (B) the content for which knowledge is estimated (quiz 1: have neither; quiz 2: have one but not the other; quiz 3: have both)
%%    \end{itemize}
%%\end{itemize}

%When we fit a GLMM to estimates of participants' knowledge for each Quiz~1
%question based on all other Quiz~1 questions, we observed an outcome consistent
%with our null hypothesis: higher estimated knowledge at the embedding
%coordinate of a held-out question was associated with a lower likelihood of
%answering the question correctly (odds ratio $(OR) = 0.136$, likelihood-ratio
%test statistic $(\lambda_{LR}) = 19.749$, 95\%\ $\textnormal{CI} = [14.352,\
%26.545],\ p < 0.001$). This outcome suggests that our knowledge estimates do
%\textit{not} provide useful information about participants' Quiz~1 performance
%when we aggregated across all question content areas. We speculated that this
%might either indicate that the knowledge estimates are uninformative in
%general, or about Quiz~1 performance in particular. This would be expected, for
%example, if participants were guessing about the answers to the Quiz~1
%questions (prior to having watched either lecture). When we repeated this
%analysis for Quizzes~2 and~3, we found that higher estimated knowledge
%for a given question predicted a \textit{greater} likelihood of answering it correctly
%(Quiz~2: $OR = 2.905,\ \lambda_{LR} = 17.333,\ 95\%\ \textnormal{CI} =
%[14.966,\ 29.309],\ p = 0.001$; Quiz~3: $OR = 3.238,\ \lambda_{LR} = 6.882,\
%95\%\ \textnormal{CI} = [6.228,\ 8.184],\ p = 0.016$). Taken together, these
%results suggest that our knowledge estimates reliably predict participants'
%performance on individual held-out quiz questions, but only after participants
%have received at least some training.

%When we fit a GLMM to estimates of participants' knowledge for each Quiz~1
%question based on all other Quiz~1 questions, we found that higher estimated
%knowledge for a given question predicted a greater likelihood of answering it 
%correctly (odds ratio $(OR) = 8.126$, 95\%\ $\textnormal{CI} = [3.116,\ 
%20.123]$, likelihood-ratio test statistic $(\lambda_{LR}) = 17.002,\ p < 
%0.001$). This relationship held and appeared to strengthen when we repeated this 
%analysis for Quiz~2 ($OR = 14.902,\ 95\%\ \textnormal{CI} = [4.976,\ 39.807],\ 
%\lambda_{LR} = 25.408,\ p < 0.001$), and again for Quiz~3 ($OR = 37.409,\ 95\%\ 
%\textnormal{CI} = [10.425,\ 107.145],\ \lambda_{LR} = 40.948,\ p < 0.001$).
%Taken together, these results suggest that our knowledge estimates reliably 
%predict participants' performance on individual questions when aggregated across
%all quiz content, and that their predictive power increases as participants gain
%additional knowledge about the content captured by the text embedding space.

When we fit a GLMM to estimates of participants' knowledge for each Quiz~1
question based on all other Quiz~1 questions, we found that higher estimated
knowledge for a given question predicted a greater likelihood of answering it 
correctly (odds ratio $(OR) = 8.126$, 95\%\ $\textnormal{CI} = [3.116,\ 
20.123]$, likelihood-ratio test statistic $(\lambda_{LR}) = 17.002,\ p < 
0.001$). This relationship held when we repeated this analysis for Quiz~2 ($OR 
= 14.902,\ 95\%\ \textnormal{CI} = [4.976,\ 39.807],\ \lambda_{LR} = 25.408,\ 
p < 0.001$) and again for Quiz~3 ($OR = 37.409,\ 95\%\ \textnormal{CI} = 
[10.425,\ 107.145],\ \lambda_{LR} = 40.948,\ p < 0.001$). Taken together, these 
results suggest that our knowledge estimates can reliably predict participants' 
performance on individual questions when aggregated across all quiz content.

We observed a similar set of results when we restricted our estimates of
participants' knowledge for questions about each lecture to consider only
their performance on other questions about the \textit{same} lecture.
Specifically, for Quiz~1, participants' knowledge of \textit{Four Fundamental 
Forces}-related questions, estimated from their performance on other 
\textit{Four Fundamental Forces}-related questions, was predictive of their 
ability to answer those questions correctly ($OR = 15.934,\ 95\%\ 
\textnormal{CI} = [5.173,\ 38.005],\ \lambda_{LR} = 40.971,\ p = 0.001$). The 
same was true of participants' estimated knowledge for \textit{Birth of 
Stars}-related questions based on their performance on other \textit{Birth of 
Stars}-related questions ($OR = 9.775,\ 95\%\ \textnormal{CI} = [2.93,\ 
25.08],\ \lambda_{LR} = 13.924,\ p = 0.001$). These within-lecture knowledge
estimates also predicted success on questions about both lectures when we 
computed them analogously for Quiz~2 (\textit{Four Fundamental Forces}: $OR = 
35.126,\ 95\%\ \textnormal{CI} = [5.113,\ 123.868],\ \lambda_{LR} = 32.251,\ 
p < 0.001$; \textit{Birth of Stars}: $OR = 4.717,\ 95\%\ \textnormal{CI} = 
[2.021,\ 9.844],\ \lambda_{LR} = 16.788,\ p < 0.001$). For Quiz~3, we found that
within-lecture knowledge estimates predicted participants' success on 
\textit{Birth of Stars}-related questions ($OR = 16.902,\ 95\%\ \textnormal{CI} 
= [3.353,\ 53.265],\ \lambda_{LR} = 23.233,\ p < 0.001$) but not on \textit{Four 
Fundamental Forces}-related questions ($OR = 2.485,\ 95\%\ \textnormal{CI} = 
[0.724,\ 8.366],\ \lambda_{LR} = 1.984,\ p = 0.170$). This may indicate that the 
within-lecture knowledge estimates are susceptible to ceiling effects in 
participants' quiz performance. On Quiz~3, after viewing both lectures, no 
participant answered more than three \textit{Four Fundamental Forces}-related 
questions incorrectly, and all but five participants (out of 50) answered two or 
fewer questions incorrectly. (This was the only subset of questions about either 
lecture, across all three quizzes, for which this was true.) Because of this, 
when we held out one incorrectly answered \textit{Four Fundamental Forces} 
question from a given participant's Quiz~3 responses and estimated their 
knowledge at its embedding coordinate using the remaining \textit{Four 
Fundamental Forces} questions they answered, for 90\% of participants, that 
estimate leveraged information about at most a single other question they were 
\textit{not} able to correctly answer. This broad homogeneity in participants' success on questions used to estimate their knowledge may have hurt our ability 
to accurately characterize the specific (and by Quiz~3, relatively few) aspects 
of the lecture content they did \textit{not} know about. Taken together, these 
results suggest that our knowledge estimates can reliably distinguish between 
questions about different content covered by a single lecture, provided there is 
sufficient diversity in participants' quiz responses to extract meaningful 
information about both what they know and what they do not know.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Finally, when we estimated participants' knowledge for each question about one
lecture using their performance on questions (from the same quiz) about the
\textit{other} lecture, we observed a somewhat different pattern of results.
Here we found that before viewing either lecture (i.e., on Quiz~1),
participants' abilities to answer \textit{Four Fundamental Forces}-related
questions could not be predicted from their responses to \textit{Birth of
Stars}-related questions ($OR = 1.896,\ 95\%\ \textnormal{CI} = [0.419,\ 
9.088],\ \lambda_{LR} = 0.712,\ p = 0.404$), nor could their abilities to 
answer \textit{Birth of Stars}-related questions be predicted from their 
responses to \textit{Four Fundamental Forces}-related questions ($OR = 1.522,\ 
95\%\ \textnormal{CI} = [0.332,\ 6.835],\ \lambda_{LR} = 0.286,\ p = 0.611$).
We similarly found that participants' success on questions about either lecture 
could not be predicted given their responses to questions about the other 
lecture after viewing \textit{Four Fundamental Forces} but before viewing \textit{Birth of Stars} (i.e., on Quiz~2; \textit{Four Fundamental Forces} 
questions given \textit{Birth of Stars} questions: $OR = 3.49,\ 95\%\ 
\textnormal{CI} = [0.739,\ 12.849],\ \lambda_{LR} = 3.266,\ p = 0.083$; 
\textit{Birth of Stars} questions given \textit{Four Fundamental Forces} 
questions: $OR = 2.199,\ 95\%\ \textnormal{CI} = [0.711,\ 5.623],\ 
\lambda_{LR} = 2.304,\ p = 0.141$). Only after viewing \textit{both} lectures 
(i.e., on Quiz~3) did these across-lecture knowledge estimates reliably predict 
participants' success on individual quiz questions (\textit{Four Fundamental 
Forces} questions given \textit{Birth of Stars} questions: $OR = 11.294,\ 
95\%\ \textnormal{CI} = [1.375,\ 47.744],\ \lambda_{LR} = 10.396,\ p < 0.001$; 
\textit{Birth of Stars} questions given \textit{Four Fundamental Forces} 
questions: $OR = 7.302,\ 95\%\ \textnormal{CI} = [1.077,\ 44.879],\ 
\lambda_{LR} = 4.708,\ p = 0.038$). Taken together, these results suggest that 
our knowledge estimates can be used to predict participants' success across 
content areas once they have received some training on both the content about 
which their knowledge is estimated, and the content used to construct these
estimates.

@@

We projected the resulting 100-dimensional topic vectors (for each second of
video and each quiz question) onto a shared 2-dimensional plane (see
\nameref{subsec:knowledge-maps}). Next, we sampled points from a $100 \times
100$ grid of coordinates that evenly tiled a rectangle enclosing the 2D
projections of the videos and questions. We used
Equation~\ref{eqn:rbf-knowledge} to estimate participants' knowledge at each of
these 10,000 sampled locations, and averaged these estimates across
participants to obtain an estimated average \textit{knowledge map}
(Fig.~\ref{fig:knowledge-maps}A). Intuitively, the knowledge map constructed
from a given quiz's responses provides a visualization of ``how much''
participants knew about any content expressible by the fitted text embedding
model at the point in time when they completed that quiz.
